---
title: "HW_3"
author: "Hector Santana"
date: "10/21/2021"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(DataExplorer)
library(knitr)
library(psych)
library(caret)
```

# HOMEWORK 3

Hector Santana, Zachary Safir, Mario Pena
 
In this assignment we will explore, analyze and build a binary logistic regression model to predict wheter a particular neighborhood will be at risk for high crime levels. 

We are provided with information on 466 neighborhoods, 12 predictor variables and 1 response variable. The response variable indicates whether the crime rate is above the median (1) or not (0).

We first download the "training" and "evaluation" datasets from GitHub:

```{r Data Importation}
training <- read.csv("https://raw.githubusercontent.com/manonfire86/Data621/main/HW_3/crime-training-data_modified.csv", header=TRUE, sep=",")

evaluation <- read.csv("https://raw.githubusercontent.com/manonfire86/Data621/main/HW_3/crime-evaluation-data_modified.csv", header=TRUE, sep=",")

training[,c("chas","target")] <- lapply(training[,c("chas","target")],factor)
```


# DATA EXPLORATION

Below we have created a table with the summary statistics of our 12 predictor variables.

```{r}
summary <- describe(training[,c(1:12)])
kable(summary)
```


According to our summary statistics above, and our graph below, there are no missing values in our dataset.

```{r}
missing_values <- training %>%
  mutate(id = row_number()) %>%
  gather(-id, key = "key", value = "val") %>%
  mutate(isna = is.na(val)) %>%
  ggplot(aes(key, id, fill = isna)) +
    geom_raster(alpha=0.8) +
    scale_fill_manual(name = "",
        values = c('steelblue', 'tomato3'),
        labels = c("Present", "Missing")) +
    labs(x = "Variable",
           y = "Row Number", title = "Missing Values in Rows") +
    coord_flip()

missing_values
```


We can also observe that the "target" variable is evenly distributed between the "1" and "0" responses.

```{r}
knitr::kable(table(training$target))
```


Let's make some obervations based on the histogram and boxplot of the variables in our data.

```{r}
plot_histogram(training)
plot_boxplot(data = training, by = "target")

```


We can also observe the correlation of our variables with eachother on the next plot.

```{r}
plot_correlation(training)
```

It seems that "target" has a strong correlation with the variables "nox", "rad", "age", "tax" and "indus".


# DATA PREPARATION


# BUILDING MODELS

Model with all predictor variables:

```{r}
model_1 <- glm(target ~., family = binomial(), data = training)
summary(model_1)
```

Model with high collinearity between response and predictor variables:

```{r}
model_2 <- glm(target ~ nox + rad + age + tax + indus, family = binomial(), data = training)
summary(model_2)
```

# SELECT MODELS
